Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-04-23 paper/graph.tex
Starting PageRank algorithm
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@583 3fc257bb-8550-0410-81b3-a5c103da83f7

\section{Graph Algoriths in MapReduce}
\label{sec:graph}

@@

It models the web as a directed graph $G(V,E)$, with each vertex $v
\in V$ representing a web page and each edge $e_{ij} \in E$
representing a hyperlink from $v_i$ to $v_j$.  The probability of
moving from $v_i$ to another vertex $v_j$ is $\alpha/d_{out}(v_i) +
(1-\alpha)/|V|$, where $\alpha$ is a user-defined parameter (usually
0.8-0.9), $d_{out}(v)$ is the outdegree of vertex $v$, and $|V|$ is
the cardinality of $V$.  The first term represents the probability of
following a given link on page $v_i$; the second represents the
probability of moving to a random page.  For pages with no outlinks,
the first term is $\alpha/|V|$, indicating equal likelihood to move to
any other page.  Equivalently, the graph can be represented by a
matrix $A$~\cite{LangvilleMeyer05a}, with matrix entries $A_{ij} =
\alpha/d_{out}(v_i)$ if vertex $v_i$ links to $v_j$.  The PageRank
algorithm, then, is simply a power-method iteration in which the
dominating computation is matrix-vector multiplication $A^T x=y$,
where $x$ is the PageRank vector from the previous iteration.

@@

$N_{\rm remain} = N_z$ \\
while $N_{\rm remain} > 0$: \\
\> Map(): \> {\bf Generate} $N_{\rm remain}/P$ random edges on each processor \\
           \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> Collate() \\
\> Reduce(): \> {\bf Remove} duplicate edges \\
              \> \> input Key = $(V_i,V_j)$, MultiValue = one or more NULLs \\
              \> \> output Key = $(V_i,V_j)$, Value = NULL \\
\> $N_{\rm remain} = N_z - N_{kv}$

>>>>

\section{Graph Algorithms in MapReduce}
\label{sec:graph}

@@

It models the web as a directed graph $G(V,E)$, with each vertex $v_i
\in V$ representing a web page and each edge $E_{ij} \in E$
representing a hyperlink from $v_i$ TO $v_j$.  The probability of
moving from $v_i$ to another vertex $v_j$ is $\alpha/d_{out}(v_i) +
(1-\alpha)/|V|$, where $\alpha$ is a user-defined parameter (usually
0.8-0.9), $d_{out}(v_i)$ is the outdegree of vertex $v_i$, and $|V|$ is
the cardinality of $V$.  The first term represents the probability of
following a given link on page $v_i$; the second represents the
probability of moving to a random page.  For pages with no outlinks,
the first term is $\alpha/|V|$, indicating equal likelihood to move to
any other page.  Equivalently, the graph can be represented by a
matrix $A$~\cite{LangvilleMeyer05a}, with matrix entries $A_{ij} =
\alpha/d_{out}(v_i)$ if vertex $v_i$ links to $v_j$.  The PageRank
algorithm, then, is essentially a power-method iteration in which the
dominating computation is matrix-vector multiplication $A^T x=y$,
where $x$ is the PageRank vector from the previous iteration.

@@

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxxx\=xxxxx\=xxxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

Input from R-MAT generator: \\
\> MapReduce object $V$ containing all graph vertices:  Key=$v_i$, Value = NULL.  \\
\> MapReduce object $E$ containing all graph edges: Key=$v_i$, Value = $v_j$. \\
Initialize MapReduce object for matrix $A$  and \\
MapReduce object $MT$ containing all vertices with no outlinks.\\
\> Copy: $A = E$ \\
\> Add: $V$ to $A$ \\
\> Collate $A$: Vertex is Key \\
\> Reduce $A$:  \\
\> \> Input:  Key = $v_i$; MultiValue = NULL,$v_j$,$v_k$,... \\
\> \> If (${nvalues} > 1$) \\
\> \> \> Compute: $d_{out}(v_i) = {nvalues} - 1$. \\
\> \> \> Output:  For each $v_j$ in MultiValue, Key = $v_i$; Value = ($v_j$, $\alpha/d_{out}(v_i)$). \\
\> \> else \\
\> \> \> Output to $MT$: Key = $v_i$; Value = 0. \\
Initialize MapReduce object for pagerank vector $x$. \\
\> Map:   \\
\> \> Input:  Graph vertices $V$. \\
\> \> Output:  For each vertex, Key = $v_i$; Value = $1/|V|$. \\
While (${residual} > {tolerance}$) \\
\> Compute contribution for random jumps and vertices with no outlinks. \\
\> \> Add: $V$ to $MT$. \\
\> \> Collate $MT$:  Vertex is Key. \\
\> \> Reduce $MT$:   \\
\> \> \> Input:  Key = $v_i$; MultiValue = $x_i, 0$ if $v_i \in {MT}$; MultiValue = $x_i$ if $v_i \notin {MT}$. \\
\> \> \> Compute:  Contributions; store in data pointer. \\
\> \> \> Output:  If ${nvalues} = 2$, Key = $v_i$; Value = 0. \\
\> \> MPI\_Allreduce with MPI\_SUM to compute global contributions. \\
\> Compute $y = A^T x$.
